---
title: "Desert Springs Data Exploration"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
runtime: shiny
---
```{r loadpackages, echo=FALSE, output=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(RODBC)
library(tidyverse)
library(shiny)
library(DT)
```

```{r functions}
selectNA <- function(data, drop.complete.cols = FALSE, cols.to.keep = NA, cols.to.ignore = NA) {
  # Given a data frame, returns only the rows that contain NA values. Optionally, omits columns that are not missing any data.
  #
  # Args:
  #   data: a data frame or tibble
  #   cols.to.ignore: a charcter vector of the names of the columns that should not be checked for missing data.
  #   drop.complete.cols: boolean value indicating whether the result should omit columns that aren't missing data.
  #   cols.to.keep: a character vector containing the names of the columns that should be returned even if drop.complete.cols is set to TRUE.
  
  # filter data to only include rows that contain one or more NA
  if (anyNA(cols.to.ignore)) {
    na.list <- data %>%
    filter_all(any_vars(is.na(.)))
  } else {
    na.list <- data %>%
    filter_at(vars(-cols.to.ignore), any_vars(is.na(.)))
  }
  
  
  # get list of columns that contain NA's
  na.cols <- na.list %>% 
    select_if(anyNA) %>%
    names()

  # if user opts to drop columns that aren't missing data, return tibble of rows and columns containing NA values, plus any additional columns specified by cols.to.keep
  if (drop.complete.cols) {
    if (anyNA(cols.to.keep)) {
      na.list <- na.list %>% 
      select(na.cols)
    } else {
      na.list <- na.list %>% 
      select(c(cols.to.keep, na.cols))
    }
  }
  
  return(na.list)
}

```

## Connecting to the database
```{r}
# DS.dsn <- 'driver={SQL Server Native Client 11.0};server=INPLAKE52V\\MOJN;database=MOJN_DS_Water;trusted_connection=Yes;applicationintent=readonly'
# DS.conn <- odbcDriverConnect(DS.dsn)

# for testing on Sarah's local SQL server instance
DS.dsn <- 'driver={SQL Server Native Client 11.0};server=INPLAKE36792JNX\\SARAH_LOCAL;database=Testing_MOJN_DS_Water;trusted_connection=Yes;applicationintent=readonly'
DS.conn <- odbcDriverConnect(DS.dsn)

# Read data from analysis views in MOJN_DS_Water database

site <- sqlFetch(DS.conn, "analysis.Site", as.is = TRUE) %>% as_tibble()
visit <- sqlFetch(DS.conn, "analysis.Visit", as.is = TRUE) %>% as_tibble()
visit.activity <- sqlFetch(DS.conn, "analysis.VisitActivity", as.is = TRUE) %>% as_tibble()
wq.DO <- sqlFetch(DS.conn, "analysis.WaterQualityDO", as.is = TRUE) %>% as_tibble()
wq.pH <- sqlFetch(DS.conn, "analysis.WaterQualitypH", as.is = TRUE) %>% as_tibble()
wq.SpCond <- sqlFetch(DS.conn, "analysis.WaterQualitySpCond", as.is = TRUE) %>% as_tibble()
wq.tempC <- sqlFetch(DS.conn, "analysis.WaterQualityTemperature", as.is = TRUE) %>% as_tibble()
discharge.estimated <- sqlFetch(DS.conn, "analysis.DischargeEstimated", as.is = TRUE) %>% as_tibble()
discharge.flow <- sqlFetch(DS.conn, "analysis.DischargeFlowCondition", as.is = TRUE) %>% as_tibble()
discharge.volumetric <- sqlFetch(DS.conn, "analysis.DischargeVolumetric", as.is = TRUE) %>% as_tibble()
disturbance <- sqlFetch(DS.conn, "analysis.Disturbance", as.is = TRUE) %>% as_tibble()
disturbance.flowmod <- sqlFetch(DS.conn, "analysis.DisturbanceFlowModification", as.is = TRUE) %>% as_tibble()
invasives <- sqlFetch(DS.conn, "analysis.Invasives", as.is = TRUE) %>% as_tibble()
riparian <- sqlFetch(DS.conn, "analysis.Riparian", as.is = TRUE) %>% as_tibble()
wildlife <- sqlFetch(DS.conn, "analysis.Wildlife", as.is = TRUE) %>% as_tibble()

# Close database connection
close(DS.conn)
```

## Exploratory analysis

### Site

#### All data
```{r site}
renderDataTable(site)
```

#### Checking for NA values

```{r}
na.site <- site %>%
  selectNA()

renderDataTable(na.site)
```
There are `r nrow(na.site)` sites that are missing data. Any rows missing data are displayed above.

#### Actively Monitored Springs

```{r}
selectInput("park", label = "Select a park to view the list of springs that are actively monitored there", choices = unique(site$ParkCode))
```

Annual springs:
```{r}
renderDataTable({
  annual <- site %>%
  filter((ParkCode == input$park) & (SampleFrame == "Annual")) %>%
  select(SiteCode, SiteName)
  annual
})
```

3-year springs:
```{r}
renderDataTable({
  three.yr <- site %>%
  filter((ParkCode == input$park) & (SampleFrame == "3Yr")) %>%
  select(SiteCode, SiteName)
  three.yr
})
```

Map:


### Visit
```{r}
renderDataTable(visit)
```

#### Checking for NA values

```{r}
na.visit <- visit %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.visit)
```
There are `r nrow(na.visit)` visits that are missing data. Any rows missing data are displayed above.


### Visit Activity
```{r}
renderDataTable(visit.activity)
```

#### Checking for NA values

```{r}
na.visit.activity <- visit.activity %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.visit.activity)
```
There are `r nrow(na.visit.activity)` visits that are missing data. Any rows missing data are displayed above.


### Water Quality - DO
```{r}
renderDataTable(wq.DO)
```

#### Checking for NA values

```{r}
na.wq.DO <- wq.DO %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"), cols.to.ignore = c("DataQualityFlagNote"))
renderDataTable(na.wq.DO)
```
There are `r nrow(na.wq.DO)` records that are missing data. Any rows missing data are displayed above.


### Water Quality - pH
```{r}
renderDataTable(wq.pH)
```

#### Checking for NA values

```{r}
na.wq.pH <- wq.pH %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"), cols.to.ignore = c("DataQualityFlagNote"))
renderDataTable(na.wq.pH)
```
There are `r nrow(na.wq.pH)` records that are missing data. Any rows missing data are displayed above.


### Water Quality - SpCond
```{r}
renderDataTable(wq.SpCond)
```

#### Checking for NA values

```{r}
na.wq.SpCond <- wq.SpCond %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"), cols.to.ignore = c("DataQualityFlagNote"))
renderDataTable(na.wq.SpCond)
```
There are `r nrow(na.wq.SpCond)` records that are missing data. Any rows missing data are displayed above.


### Water Quality - Temperature
```{r}
renderDataTable(wq.tempC)
```

#### Checking for NA values

```{r}
na.wq.tempC <- wq.tempC %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"), cols.to.ignore = c("DataQualityFlagNote"))
renderDataTable(na.wq.tempC)
```
There are `r nrow(na.wq.tempC)` records that are missing data. Any rows missing data are displayed above.


### Discharge - Estimated
```{r}
renderDataTable(discharge.estimated)
```

#### Checking for NA values

```{r}
na.discharge.est <- discharge.estimated %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"), cols.to.ignore = c("DischargeFlag"))
renderDataTable(na.discharge.est)
```
There are `r nrow(na.discharge.est)` records that are missing data. Any rows missing data are displayed above.


### Discharge - Volumetric
```{r}
renderDataTable(discharge.volumetric)
```

#### Checking for NA values

```{r}
na.discharge.est <- discharge.estimated %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"), cols.to.ignore = c("DischargeFlag"))
renderDataTable(na.discharge.est)
```
There are `r nrow(na.discharge.est)` records that are missing data. Any rows missing data are displayed above.

### Discharge - Flow
```{r}
renderDataTable(discharge.flow)
```

#### Checking for NA values

```{r}
na.discharge.flow <- discharge.flow %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.discharge.flow)
```
There are `r nrow(na.discharge.flow)` records that are missing data. Any rows missing data are displayed above.


### Disturbance
```{r}
renderDataTable(disturbance)
```

#### Checking for NA values

```{r}
na.disturbance <- disturbance %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.disturbance)
```
There are `r nrow(na.disturbance)` records that are missing data. Any rows missing data are displayed above.


### Disturbance - Flow Modifications
```{r}
renderDataTable(disturbance.flowmod)
```

#### Checking for NA values

```{r}
na.disturbance.flowmod <- disturbance.flowmod %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.disturbance.flowmod)
```
There are `r nrow(na.disturbance.flowmod)` records that are missing data. Any rows missing data are displayed above.


### Invasives
```{r}
renderDataTable(invasives)
```

#### Checking for NA values

```{r}
na.invasives <- invasives %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.invasives)
```
There are `r nrow(na.invasives)` records that are missing data. Any rows missing data are displayed above.


### Riparian Veg
```{r}
renderDataTable(riparian)
```

#### Checking for NA values

```{r}
na.riparian <- riparian %>%
  selectNA(drop.complete.cols = TRUE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.riparian)
```
There are `r nrow(na.riparian)` records that are missing data. Any rows missing data are displayed above.


### Wildilfe
```{r}
renderDataTable(wildlife)
```

#### Checking for NA values

```{r}
na.wildlife <- wildlife %>%
  selectNA(drop.complete.cols = FALSE, cols.to.keep = c("ParkCode", "SiteCode", "SiteName", "VisitDate"))
renderDataTable(na.wildlife)
```
There are `r nrow(na.wildlife)` records that are missing data. Any rows missing data are displayed above.

